---
title: "02_21_Nightbreak"
author: "Lin_Zhang"
date: "2/21/2020"
output: 
  html_document: 
    keep_md: yes
---
```{r}
library(rethinking)
library(dplyr)
```
Q1: a. Load the dataset. Look for column "Score" for the response variable we are interested in. A developmental score of 1 or 2 indicates vegetative growth, while a score of 3,4 or 5 indicates reproductive growth. Create a "Reproduction" column with values 0 and 1, where 0 indicates vegetative growth and 1 indicates reproductive growth. 
```{r}
nightbreak <- read.csv(file = "Nightbreak_02_08_20_Rclub.csv",header = TRUE,sep = ",",stringsAsFactors = FALSE)
d <- nightbreak
```

b. (optional) Take a look at columns "loc1" through "loc5". The levels "A" or "P" indicate which parental allele the RIL has inherited. Can you think of a reason why there are 32 lines in this experiment?
5 loci,each with 2 allele options, 2^5

Q2:
Using the "Reproduction" column you created in Q1a as the response variable, fit a simple model with effects of genotype ("RIL") and treatment ("Treatment") on the rate of transitioning to reproductive growth by 2/8/20. (Things you might want to consider: appropriate likelihood function, intersection term). 
```{r error= TRUE}
d$Reproduction <- if_else(d$Score < 3,0L,1L)
d$nightbreak <- if_else(d$Treatment=="NightBreak",1L,0L)
d$curtain <- if_else(d$Treatment=="Control_NoCurtain",0L,1L)

dat <- list(
  NB = d$nightbreak,
  CR = d$curtain,
  G = as.numeric(as.factor(d$RIL)),
  R = d$Reproduction
)

m1 <- ulam(
  alist(
    R ~ dbinom(1,p),
    logit(p) <- a[G]+bN*NB+bC*CR+INC[G]*CR+INB[G]*NB,
    a[G] ~ dnorm(0,1.5),  # inv_logit(-3) to inv_logit(3)
    c(bN,bC) ~ dnorm(0,1.5),
    INC[G] ~ dnorm(0,.5),
    INB[G] ~ dnorm(0,.5)
    ), data = dat, chains = 4, log_lik = TRUE
)
# why cannot go through after adding interaction
```

```{r}
traceplot(m1)
precis(m1,depth = 2)
```

```{r}
coef(m1)["a"]
inv_logit(coef(m1)["a"]) # flowering coefficient for control wo curtain, wo nightbreak
summary()  #why higher here
```

```{r}
inv_logit(coef(m1)["a"]+coef(m1)["b_nb"]) # coefficient for adding night break
```

Q3:
Because we are more interested in the effects of individual loci than the performance of specific genotypes, fit a model with additive effects of the five loci and effect of treatment.  
```{r}

```

5 loci responsible for daylength from QTL 
loci 2,3,5 promote flowering, from parent P
however, no interaction seen between 5 loci and the treatment
nightbreak would promote flowering regardless of the loci
the loci would promote flowering regardless of nightbreak

How long is the night length in the nightbreak and field experiment?
Greenhouse warmer than the field

Q5 interesting by ordered categorical outcome
Q4 model: anyway to simplify ??


the loci not general regulator of development (1,2,3,4,5) but regulator of reproduction

Pooling across plots (autocorrelation spatial relationship)
Pooling across replicates
Puting two controls together

Q7.
nightbreak increases reproduction probability 
A is the one suppose to be responsing to nighbreak , A is the one needs long days to flower

plot flowering time ~ genotype, to see if the model are making sense of the raw data

flip loci coding -> difference? (maybe due to logit transform)

Phenotype: first time to flowering (scoring every day)  --- right now the proportion data more look like my sporulation percentage

Also segregate for height, this population
Prior prediction, function
posterior plot

? censored cat problem ?

# apply to my data
0,5,10,50,200 ppm treatment (ordered categorical predictor variable), 
Response: sporulation percentage (sporulate percentage out of 10 seedlings, Poisson with a fixed interval),
Timepoints: 7,11,15 dpi  (time-course, censored cat?)
isolate ~ genotype as index variable
plant cultured on
no rep, no plot

Question:
genotype ~ treatment interaction
Poisson

Python for datascience: not very well but know how to find out
Pandas, Nonpy

Covariance between parameters, covariance is a pain? covariance in linear regression
14.1
Video  

