---
title: "02_21_Nightbreak"
author: "Lin_Zhang"
date: "2/21/2020"
output: 
  html_document: 
    keep_md: yes
---
```{r}
library(rethinking)
```
Q1: a. Load the dataset. Look for column "Score" for the response variable we are interested in. A developmental score of 1 or 2 indicates vegetative growth, while a score of 3, 4, or 5 indicates reproductive growth. Create a "Reproduction" column with values 0 and 1, where 0 indicates vegetative growth and 1 indicates reproductive growth. 
```{r}
nightbreak <- read.csv(file = "Nightbreak_02_08_20_Rclub.csv",header = TRUE,sep = ",",stringsAsFactors = FALSE)
nightbreak$Reproduction <- ifelse(nightbreak < 4, 0,1)
```

b. (optional) Take a look at columns "loc1" through "loc5". The levels "A" or "P" indicate which parental allele the RIL has inherited. Can you think of a reason why there are 32 lines in this experiment?
5 loci,each with 2 allele options, 2^5

Q2:
Using the "Reproduction" column you created in Q1a as the response variable, fit a simple model with effects of genotype ("RIL") and treatment ("Treatment") on the rate of transitioning to reproductive growth by 2/8/20. (Things you might want to consider: appropriate likelihood function, intersection term). 
```{r error= TRUE}
d <- nightbreak
d$Treatment <- as.factor(d$Treatment)
treat_levels <- c(1:3)
d$treat_levels <- treat_levels[d$Treatment]
unique(d$RIL)
d$RIL[d$RIL=="PI"] <- c("0")
d$RIL[d$RIL=="Arm"] <- c("1")


dat <- list(
  TR = d$treat_levels,
  G = as.numeric(d$RIL),
  R = d$Reproduction
)

m1 <- ulam(
  alist(
    R ~ dbinom(1,p),
    logit(p) <- a[G] + bT*TR,
    a[G] ~ dnorm(0,1.5),
    bT ~ dnorm(0,1.5)
    ), data = dat, chains = 4, log_lik = TRUE
)
```

Q3:
Because we are more interested in the effects of individual loci than the performance of specific genotypes, fit a model with additive effects of the five loci and effect of treatment.  

