---
title: "Chapter9_HW"
author: "Lin Zhang"
date: "9/27/2019"
output: 
  html_document: 
    keep_md: yes
---
##8E1.Which of the following is a requirement of the simple Metropolis algorithm? 
#(1) The parameters must be discrete. 
#(2) The likelihood function must be Gaussian. 
#(3) The proposal distribution must be symmetric.
## Answer:(3)

##8E2.Gibbs sampling is more efficient than the Metropolis algorithm. How does it achieve this extra efficiency? Are there any limitations to the Gibbs sampling strategy?
## Answer: using adaptive proposal,in which distribution of proposed parameter values adjusts itself intelligently,depending upon the parameter values at the moment. The adaptive proposals depends upon using particular combinations of prior distributions and likelihoods known as conjugate pairs.(not clear about the detail)

##8E3.Which sort of parameters can Hamiltonian Monte Carlo not handle? Can you explain why?
## Answer: HMC cannot deal with discrete parameters.Because it needs to calculate gradient, i.e. local curvature at each location, i.e. current vector of parameter values.

##8M1.Re-estimate the terrain ruggedness model from the chapter, but nowusing a uniform prior and an exponential prior for the standard deviation, sigma. The uniform prior should be dunif(0,10) and the exponential should be dexp(1). Do the different priors have any detectible influence on the posterior distribution?

```{r}
library(rethinking)
```

```{r original prior}
m9.2 <- map2stan(   #why substitute ulam for map2stan?
  alist(
    log_gdp_std ~ dnorm(mu,sigma),
    mu <- a[cid]+b[cid]*(rugged_std-0.215),
    a[cid] ~ dnorm(1,0.1),
    b[cid] ~ dnorm(0,0.3),
    sigma ~ dexp(1)
  ),data = dat_slim,chains = 2,cores = 2,iter = 1000
)
```

```{r}
precis(m9.2,2)
```

```{r modified prior}
library(rethinking)
m_8m1 <- map2stan(
  alist(
    log_gdp_std ~ dnorm(mu,sigma),
    mu <- a[cid]+b[cid]*(rugged_std-0.215),
    a[cid] ~ dunif(0,10),
    b[cid] ~ dunif(0,10),
    sigma ~ dexp(1)
  ),data = dat_slim,chains = 2,cores = 2,iter = 1500,control = list(adapt_delta=0.99)
)
```
```{r}
pairs(m_8m1)
precis(m_8m1,2)
```


# the posterior distribution of b[2], i.e. non-African slope change to nearly flat as shown below, but surprisingly all other parameters reach a identical estimation, understand priors better
```{r}
#when use ulam, it works as below, the esimate is nearly the same as using map2stan
precis(m_8m1,2)
show(m_8m1)
traceplot(m_8m1)
trankplot(m_8m1) #function not found
```

##8M2.
```{r}

```

##8H1
```{r}
mp <- map2stan(
  alist(
    a ~ dnorm(0,1),
    b ~ dcauchy(0,1)
  ),
  data = list(y=1),
  start = list(a=0,b=0),
  iter = 1e4,warmup = 300,WAIC = FALSE,control = list(max_treedepth=15)
)
```
```{r}
precis(mp)
pairs(mp)
```

##8H2.Recall the divorce rate example from Chapter 5. Repeat that analysis, using map2stan this time, fitting models m5.1, m5.2, and m5.3. Use compare to compare the models on the basis of WAIC. Explain the results
```{r}
# load data and copy
library(rethinking)
data("WaffleDivorce")
d <- WaffleDivorce
```

```{r}
d$A <- scale(d$MedianAgeMarriage)
d$D <- scale(d$Divorce)
```

```{r}
dat_slim <- list(
  Divorce_std=d$D,
  MedianAge_std=d$A
)
str(dat_slim)
```


```{r}
m5.1 <- map2stan(  
  alist(
    Divorce_std ~ dnorm(mu,sigma),
    mu <- a+bA*MedianAge_std,
    a ~ dnorm(0,0.2),
    bA ~ dnorm(0,0.5),
    sigma ~ dexp(1)
  ),data = dat_slim
)
```

```{r}
precis(m5.1)
pairs(m5.1)
```

```{r}
d$M <- scale(d$Marriage)
dat_slim <- list(
  Divorce_std=d$D,
  Marriage_rate_std=d$M
)
str(dat_slim)
```

```{r}
m5.2 <- map2stan(
  alist(
    Divorce_std ~ dnorm(mu,sigma),
    mu <- a+bM*Marriage_rate_std,
    a ~ dnorm(0,0.2),
    bM ~ dnorm(0,0.5),
    sigma ~ dexp(1)
  ),data=dat_slim
)
```

```{r}
precis(m5.2)
pairs(m5.2)
```

```{r}
dat_slim <- list(
  Divorce_std=d$D,
  MedianAge_std=d$A,
  Marriage_rate_std=d$M
)
str(dat_slim)
```

```{r}
m5.3 <- map2stan(
  alist(
    D ~ dnorm(mu,sigma),
    mu <- a+bA*MedianAge_std+bM*Marriage_rate_std,
    a ~ dnorm(0,0.2),
    bM ~ dnorm(0,0.5),
    bA ~ dnorm(0,0.5),
    sigma ~ dexp(1)
  ),data = dat_slim
)
```

```{r}
precis(m5.3,2)
pairs(m5.3)
```

```{r}
compare(m5.1,m5.2,m5.3)
```

