---
title: "Chapter_10_HW"
author: "Lin Zhang"
date: "11/3/2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
library(rethinking)
library(tidyverse)
```

```{r data import 01}
data("Wines2012") # no replicates in the dataset
head(Wines2012)
str(Wines2012)
summary(Wines2012$score)
d <- Wines2012
```

```{r data examine 01}
unique(d$judge)
unique(d$wine)
summary(d$score)
```
# model score by variation among judges and wines
## use ratio agianst max as scale for score, cannot converge MCMC!
```{r data standardization 01}
# sapply(d[,c("score","judge","wine")],function(x)sum(is.na(x))) # no missing data in the three variables
# d$S <- d$score/max(d$score) # not well mixed chain generated, also when trying constructing two models, the second one cannot run through with prompt of "recompile to avoid R session crash"
# d$judge_id <- as.integer(d$judge)
# #d$J <- scale(d$judege_id)
# d$wine_id <- as.integer(d$wine)
# #d$W <- scale(d$wine_id)
```

```{r DAG plot 01}
library(dagitty)
dag <- dagitty("dag{
J->S
W->S
}")
coordinates(dag) <- list(x=c(W=0,J=2,S=1),y=c(S=0,W=0,J=0))
drawdag(dag)
```

## use Z-score as scale for score
```{r scale S use scale 01}
sapply(d[,c("score","judge","wine")],function(x)sum(is.na(x))) # no missing data in the three variables
d$S <- scale(d$score) # why use scale here
d$judge_id <- as.integer(d$judge)
#d$J <- scale(d$judege_id)]
d$wine_id <- as.integer(d$wine)
#d$W <- scale(d$wine_id)
```

```{r construct a clean data list Rcode 9.11 01}
dat_slim <- list(
  score_std = d$S,
  wine = as.integer(d$wine_id),
  judge = as.integer(d$judge_id)
)
str(dat_slim)
```

```{r build model for score using two index variable wine_id and judge_id Rcode 8.13 01}
# S ~ Normal(u,sigma), ui=a[i]+beta_w(w-w_mean)+beta_judge(j-j_mean)
# why use mcmc is better here
m1 <- ulam(alist(
    score_std ~ dnorm(mu,sigma),
    mu <- a_w[wine]+a_J[judge],  #interaction between categorical variables is just additive?
    a_w[wine] ~ dnorm(0,1),
    a_J[judge] ~ dnorm(0,1),
    sigma ~ dexp(1)),
  data = dat_slim, 
  chains = 4,
  cores = 4, 
  iter = 2000  # iteration time setting?
)
# if not run with a fresh R session, prompt recompiling to aviod R session crash
```

```{r 01 Diagnostics_01}
show(m1)  
precis(m1,2)  #n_eff is ~1000-2000 for a compared to 4000 total sample number
plot(precis(m1,2)) 
# interpret variation among individual judges and individual wines
# variation among judges is bigger than that among wines, might due to original data structure, each judge is examing 20 different wines
# plot the differences?
# Judge 5 gives the highest rating, Judge 8 gives the lowest rating
# Wine 4 was rated best on average, wine 18 was justed worst on average
```

```{r 01 Diagnostics_02,fig.height=20,fig.width=20}
pairs(m1)
```

```{r 01 Diagnostics_03}
traceplot(m1)
```

```{r 01 Diagnostics_04}
trankplot(m1)
```

```{r 01 extract prior}
set.seed(7)
prior <- extract.prior(m1)
str(prior)
```
```{r 01 plot prior}
data <- data.frame(judge=sample(1:9,1000,replace = TRUE),
                   wine=sample(1:20,1000,replace = TRUE))
prior.score <- link(m1,data,post=prior)
dens(prior.score)
dens(dat_slim$score) # a bit weird
```

```{r 01 adopted from chapter 8 tulip example} 
##???
par(mfrow=c(3,3))
for (j in 1:9) {
  idx <- which(dat_slim$judge==j) #fixed judge
  plot(dat_slim$wine[idx],dat_slim$score_std[idx],xlim=c(1,20),ylim=c(0.3,1),
       xlab="wine",ylab="score",pch=16,col=rangi2)
  mu <- link(m1,post=prior,data = data.frame(judge=j,wine=1:20))
  for(i in 1:20)lines(1:20,mu[i,],col=col.alpha("blue",0.3))
}
```

## 2. consider three features of wines and judges
## don't include interaction effects, but model based on flight, wine.amer, judge.amer, omit individual judge and wine index variables
```{r 02 data import}
data("Wines2012") # no replicates in the dataset
head(Wines2012)
str(Wines2012)
summary(Wines2012$score)
d <- Wines2012
```

```{r 02 data examine}
unique(d$judge.amer)
unique(d$wine.amer)
unique(d$flight)
summary(d$score)
```

## use Z-score as scale for score
```{r 02 scale S use scale}
sapply(d[,c("score","judge","wine","flight","wine.amer","judge.amer")],function(x)sum(is.na(x))) # no missing data in the three variables
d$S <- scale(d$score) # why use scale here
d$judge_amer_id <- as.integer(d$judge.amer)
#d$J <- scale(d$judege_id)]
d$wine_amer_id <- as.integer(d$wine.amer)
d$flight_id <- as.integer(d$flight)
unique(d$flight_id)
unique(d$judge_amer_id)
unique(d$wine_amer_id)
#d$W <- scale(d$wine_id)

```

```{r 02 construct a clean data list Rcode 9.11}
dat_slim <- list(
  score_std = d$S,
  wine_id = as.integer(d$wine_amer_id)+1, # why 0 1 cannot go through ulam model below
  judge_id = as.integer(d$judge_amer_id)+1,
  flight_id = as.integer(d$flight_id)
)
str(dat_slim)
```

```{r 02 build model for score using two index variable wine_id and judge_id Rcode 8.13}
# S ~ Normal(u,sigma), ui=a[i]+beta_w(w-w_mean)+beta_judge(j-j_mean)
# why use mcmc is better here
m2 <- ulam(alist(
    score_std ~ dnorm(mu,sigma),
    mu <- a_w[wine_id]+a_J[judge_id]+a_f[flight_id],  #interaction between categorical variables is just additive?
    a_w[wine_id] ~ dnorm(0,1),
    a_J[judge_id] ~ dnorm(0,1),
    a_f[flight_id] ~ dnorm(0,1),
    sigma ~ dexp(1)),
  data = dat_slim, 
  chains = 4,
  cores = 4, 
  iter = 2000  # iteration time setting?
)
# if not run with a fresh R session, prompt recompiling to aviod R session crash
```


```{r 02 extract prior}
set.seed(7)
prior <- extract.prior(m2)
str(prior)
```

```{r 02 plot prior}
data <- data.frame(judge_id=sample(0:1,1000,replace = TRUE)+1,
                   wine_id=sample(0:1,1000,replace = TRUE)+1,
                   flight_id=sample(1:2,1000,replace = TRUE))
prior.score <- link(m2,data,post=prior)
dens(prior.score) # looks alright
dens(dat_slim$score) # a bit weird
```

```{r 02 check the chains}
show(m2)
precis(m2,2)
plot(precis(m2,2))
# differences between wines and judges
# American wines have slightly lower score than French, American judges gave higher score than French judges, differences are minor
# difference among judge categories are bigger than that of wine origin catogires
# wine catogries, i.e. either white or red wine does not influence the scores of wine
# Related to results in Problem1, differences in judge or wine is not due to wine flight or country/origin of judge/wine
```


```{r 02 Diagnostics_02,fig.height=20,fig.width=20}
pairs(m2)
```

```{r 02 Diagnostics_03}
traceplot(m2)
```

```{r 02 Diagnostics_04}
trankplot(m2)
```

## 3. now consider two-way interactions among the three features

```{r 03 data import}
rm(list = ls(all.names = TRUE))
data("Wines2012") # no replicates in the dataset
head(Wines2012)
str(Wines2012)
summary(Wines2012$score)
d <- Wines2012
```

```{r 03 data examine}
unique(d$judge.amer)
unique(d$wine.amer)
unique(d$flight)
summary(d$score)
```

## use Z-score as scale for score
```{r 03 scale S use scale}
sapply(d[,c("score","judge","wine","flight","wine.amer","judge.amer")],function(x)sum(is.na(x))) # no missing data in the three variables
d$S <- scale(d$score) # why use scale here
d$judge_amer_id <- as.integer(d$judge.amer)
#d$J <- scale(d$judege_id)]
d$wine_amer_id <- as.integer(d$wine.amer)
d$flight_id <- as.integer(d$flight)
unique(d$flight_id)
unique(d$judge_amer_id)
unique(d$wine_amer_id)
#d$W <- scale(d$wine_id)

```

```{r 03 construct a clean data list Rcode 9.11}
dat_slim <- list(
  score_std = d$S,
  wine_id = as.integer(d$wine_amer_id)+1, # why 0 1 cannot go through ulam model below
  judge_id = as.integer(d$judge_amer_id)+1,
  flight_id = as.integer(d$flight_id)
)
str(dat_slim)
```

```{r 03 build model for score using two index variable wine_id and judge_id Rcode 8.13}
# S ~ Normal(u,sigma), ui=a[i]+beta_w(w-w_mean)+beta_judge(j-j_mean)
# why use mcmc is better here
m3 <- ulam(alist(
    score_std ~ dnorm(mu,sigma),
    mu <- a_w[wine_id]+
          a_J[judge_id]+
          a_f[flight_id]+
          i_f_w[flight_id]*(wine_id-1)+ # interaction between flight and wine origin, #1 if american wine, 0 if french wine, assume there is more uncertainty in American wine?  # as seen in Problem 2, no difference in wine flight, why here examine interaction of wine flight with others
          i_f_j[flight_id]*(judge_id-1)+ # interaction between flight and judge country,# 1 if american judge, 0 if french judge, assume there is more uncertainty in American judge?
          i_w_j[wine_id]*(judge_id-1), # interaction between wine origin and judge country, # 1 of american judge, 0 if french judge, assume there is more uncertainty in American judge?
    a_w[wine_id] ~ dnorm(0,1),
    a_J[judge_id] ~ dnorm(0,1),
    a_f[flight_id] ~ dnorm(0,1),
    i_f_w[flight_id] ~ dnorm(0,.5),# what is the meaning of [flight_id] in this parameter?? 
    i_f_j[flight_id] ~ dnorm(0,.5),
    i_w_j[wine_id] ~ dnorm(0,.5),
    sigma ~ dexp(1)),
  data = dat_slim, 
  chains = 4,
  cores = 4, 
  iter = 2000  # iteration time setting?
)
# if not run with a fresh R session, prompt recompiling to aviod R session crash
```


```{r 03 extract prior}
set.seed(7)
prior <- extract.prior(m3)
str(prior)
```

```{r 03 plot prior}
data <- data.frame(judge_id=sample(0:1,1000,replace = TRUE)+1,
                   wine_id=sample(0:1,1000,replace = TRUE)+1,
                   flight_id=sample(1:2,1000,replace = TRUE))
prior.score <- link(m3,data,post=prior)
dens(prior.score) # looks alright, change bandwidth?
dens(dat_slim$score) # a bit weird
```

```{r 03 check the chains}
show(m3)
precis(m3,2)
plot(precis(m3,2))
##???
# white wines has a larger interaction with wine origin, i.e. American white wine and Frech white wine change the score !!
```

```{r 03 Diagnostics_02,fig.height=20,fig.width=20}
pairs(m3)
```

```{r 03 Diagnostics_03}
traceplot(m3)
```

```{r 03 Diagnostics_04}
trankplot(m3)
```


