---
title: "Chapter_13_HW"
output: 
  html_document: 
    keep_md: yes
---
#12E1. Which of the following priors will produce more shrinkage in the estimates? (a) αtank ∼
Normal(0,1); (b) αtank ∼ Normal(0,2).
(a),as this prior is more narrow and produce better prediction about the population mean cross clusters (tanks) data.

#12E2.Make the following model into a multilevel model.
yi ∼ Binomial(1,pi)
logit(pi) = αgroup[i] + βxi
αgroup ∼ Normal(0,10)
β ∼ Normal(0,1)

yi ∼ Binomial(1,pi)
logit(pi) = αgroup[i] + βxi 
agroup ~ Normal(a_bar,sigma)
a_bar ~ Normal(0,1.5)
sigma ~ Exponential (1) #why exponential

#12E3.Make the following model into a multilevel model.
yi ∼ Normal(µi,σ)
µi = αgroup[i] + βxi
αgroup ∼ Normal(0,10)
β ∼ Normal(0,1)
σ ∼ HalfCauchy(0,2)

yi ∼ Normal(µi,σ)
µi = αgroup[i] + βxi
αgroup ∼ Normal(a_bar,sigma)
a_bar ~ Normal(0,1.5)
sigma ~ Exponential
β ∼ Normal(0,1)
σ ∼ HalfCauchy(0,2)

#12M1.Revisit the Reed frog survival data, data(reedfrogs), and add the predation and size
treatment variables to the varying intercepts model. Consider models with either main eﬀect alone,
both main eﬀects, as well as a model including both and their interaction. Instead of focusing on
inferences about these two predictor variables, focus on the inferred variation across tanks. Explain why it changes as it does across models.

```{r}
data(reedfrogs) 
d <- reedfrogs
str(d)
```

```{r}
d$tank <- 1:nrow(d)
d$pred_id <- ifelse(d$pred=="no",0L,1L)
d$size_id <- ifelse(d$size=="small",0L,1L)
```

## model with only predation as main effect
```{r}
dat <- list(S=d$surv,
            P=d$pred_id,
            tank=d$tank,
            N=d$density) 
  
m12M1.1 <- ulam(
  alist(
    S ~ dbinom(N,p),
    logit(p) <- a[tank]+bpred*P,
    a[tank] ~ dnorm(a_bar,sigma),
    bpred ~ dnorm(0,0.3),
    a_bar ~ dnorm(0,1.5),
    sigma ~ dexp(1)
  ),data = dat, chains = 4, log_lik = TRUE,iter = 4000
)
```
```{r}
precis(m12M1.1,depth = 2)
summary(precis(m12M1.1,depth = 2)$mean)
```

## model with only size as main effect
```{r}
dat <- list(S=d$surv,
            Z=d$size_id,
            tank=d$tank,
            N=d$density) 
  
m12M1.2 <- ulam(
  alist(
    S ~ dbinom(N,p),
    logit(p) <- a[tank]+bsize*Z,
    a[tank] ~ dnorm(a_bar,sigma),
    bsize ~ dnorm(0,0.3),
    a_bar ~ dnorm(0,1.5),
    sigma ~ dexp(1)
  ),data = dat, chains = 4, log_lik = TRUE,iter = 2000
)
```
```{r}
precis(m12M1.2,depth = 2)
boxplot(precis(m12M1.1,depth = 2)$mean,precis(m12M1.2,depth = 2)$mean)
```
more variation with only size as predictor variable

```{r}
compare(m12M1.1,m12M1.2)
```
mode with only predation as predicator varaible is better at prediction. And it expect less variation in varying intercept. Predation is more important as a predictor variable.

## model with both predation and size as main effect
```{r}
dat <- list(S=d$surv,
            Z=d$size_id,
            P=d$pred_id,
            tank=d$tank,
            N=d$density) 
  
m12M1.3 <- ulam(
  alist(
    S ~ dbinom(N,p),
    logit(p) <- a[tank]+bsize*Z+bpred*P,
    a[tank] ~ dnorm(a_bar,sigma),
    c(bsize,bpred) ~ dnorm(0,0.3),
    a_bar ~ dnorm(0,1.5),
    sigma ~ dexp(1)
  ),data = dat, chains = 4, log_lik = TRUE,iter = 2000
)
```

```{r}
precis(m12M1.3,depth = 2)
boxplot(precis(m12M1.1,depth = 2)$mean,precis(m12M1.2,depth = 2)$mean,precis(m12M1.3,depth = 2)$mean)
```
Variation across tanks is similar to model with only predation as predictor

## model with both and interaction 
```{r}
dat <- list(S=d$surv,
            Z=d$size_id,
            P=d$pred_id,
            tank=d$tank,
            N=d$density) 
  
m12M1.4 <- ulam(
  alist(
    S ~ dbinom(N,p),
    logit(p) <- a[tank]+bsize*Z+bpred*P+bsp*Z*P,
    a[tank] ~ dnorm(a_bar,sigma),
    c(bsize,bpred,bsp) ~ dnorm(0,0.3),
    a_bar ~ dnorm(0,1.5),
    sigma ~ dexp(1)
  ),data = dat, chains = 4, log_lik = TRUE,iter = 2000
)
```
```{r}
precis(m12M1.4,depth = 2)
boxplot(precis(m12M1.1,depth = 2)$mean,precis(m12M1.2,depth = 2)$mean,precis(m12M1.3,depth = 2)$mean,precis(m12M1.4,depth = 2)$mean)
```
The variance is even smaller between tanks. So predation is the major predictor variable, with some interaction with size of tadpoles.


#12M2.Compare the models you ft just above, using WAIC. Can you reconcile the diﬀerences in
WAIC with the posterior distributions of the models?
```{r}
compare(m12M1.1,m12M1.2,m12M1.3,m12M1.4)
```
Predation-only model is the best, and very similar to predation,size-both and interaction-included model.


#12H1.In 1980, a typical Bengali woman could have 5 or more children in her lifetime. By the
year 200, a typical Bengali woman had only 2 or 3. You’re going to look at a historical set of data,when contraception was widely available but many families chose not to use it. Tese data reside in data(bangladesh) and come from the 1988 Bangladesh Fertility Survey. Each row is one of 1934 women. Tere are six variables, but you can focus on three of them for this practice problem:
(1) district: ID number of administrative district each woman resided in
(2) use.contraception: An indicator (0/1) of whether the woman was using contraception
(3) urban: An indicator (0/1) of whether the woman lived in a city, as opposed to living in a
rural area
Te frst thing to do is ensure that the cluster variable, district, is a contiguous set of integers. Recall that these values will be index values inside the model. If there are gaps, you’ll have parameters for
which there is no data to inform them. Worse, the model probably won’t run. Look at the unique
values of the district variable:
```{r}
data("bangladesh")
d <- bangladesh
```

```{r}
sort(unique(d$district))
```
```{r}
d$district_id <- as.integer(as.factor(d$district))
sort(unique(d$district_id))
```

Now, focus on predicting use.contraception, clustered by district_id. Do not include
urban just yet. Fit both (1) a traditional fxed-eﬀects model that uses dummy variables for district and (2) a multilevel model with varying intercepts for district. Plot the predicted proportions of women in each district using contraception, for both the fxed-eﬀects model and the varying-eﬀects model. That is, make a plot in which district ID is on the horizontal axis and expected proportion using contraception is on the vertical. Make one plot for each model, or layer them on the same plot, as you prefer. How do the models disagree? Can you explain the pattern of disagreement? In particular, can you explain the most extreme cases of disagreement, both why they happen where they do and why the models reach diﬀerent inferences?

predicting use.contraception, clustered by district_id. 
(1)traditional fxed-eﬀects model that uses dummy variables for district
```{r}
Density <- as.data.frame(table(d$district_id))
use_contrac_aggre <- aggregate(d$use.contraception,list(district=d$district_id),sum)
```

```{r}
dat <- list(
  C = use_contrac_aggre$x,
  district = c(1:60),
  N = Density$Freq
)

m12H1.1 <- ulam(
  alist(
    C ~ dbinom(N,p),
    logit(p) <- a[district],
    a[district] ~ dnorm(0,1.5)
  ),data = dat,chains = 4,log_lik = TRUE
)
```
(2) a multilevel model with varying intercepts for district
```{r}

```

12M3
12H2