---
title: "Chap16_HW"
author: "Lin_Zhang"
date: "8/2/2020"
output: 
  html_document: 
    keep_md: yes
---

16M1. Modify the cylinder height model, m16.1, so that the exponent 3 on height is instead a free
parameter. Do you recover the value of 3 or not? Plot the posterior predictions for the new model.
How do they differ from those of m16.1?

```{r}
library(rethinking)
data(Howell1)
d <- Howell1

# scale observed variables
d$w <- d$weight / mean(d$weight)
d$h <- d$height / mean(d$height)
```

```{r}
m16M1 <- ulam(
  alist(
    w ~ dlnorm(mu,sigma),
    exp(mu) <- 3.141593 * k *p^2 * h^beta,
    p ~ beta(2,18),
    k ~ exponential(0.5),
    beta ~ exponential(1),
    sigma ~ exponential(1)
  ),data = d,chains = 4,cores = 4,control = list(max_treedepth = 15)
)
```
```{r}
precis(m16M1)
plot(precis(m16M1))
```
not 3, probably mean the cylinder model is over-simplified,but nevertheless it is good to start with.
p,k stay very similar while even smaller variation in sigma.

```{r}
pairs(m16M1)
```
```{r}
# plot posterior distribution m16M1
h_seq <- seq( from=0 , to=max(d$h) , length.out=30 )
w_sim <- sim( m16M1 , data=list(h=h_seq) )
mu_mean <- apply( w_sim , 2 , mean )
w_CI <- apply( w_sim , 2 , PI )
plot( d$h , d$w , xlim=c(0,max(d$h)) , ylim=c(0,max(d$w)) , col=rangi2 ,
lwd=2 , xlab="height (scaled)" , ylab="weight (scaled)" )
lines( h_seq , mu_mean )
shade( w_CI , h_seq )
```
Fits better to lower height,capture the change in sub-structure of the data, ex. children/age here 
`{r}
curve( dnorm( x , 178 , 20 ) , from=100 , to=250 )
16M2. Conduct a prior predictive simulation for the cylinder height model. Begin with the priors
in the chapter. Do these produce reasonable prior height(does it mean weight here?) distributions? If not, which modifications do you suggest?

```{r}
set.seed(7)
prior <- extract.prior(m16M1)
```
```{r}
range(d$h)
```
```{r}
range(d$w)
```
```{r}
plot( NULL , xlim=range(d$h) , ylim=range(d$w) , xlab="height" , ylab="weight")
mu <- link( m16M1 , post = prior, data=list(h=range(d$h) ) )
for ( i in 1:1000) lines( c(0,1.5) , mu[i,] , col=col.alpha("black",0.3) )
```
Another way to plot the prior distribution but not sure why it doesn't work
```{r}
set.seed(2971)
N <- 100
p <- rbeta(N,2,18)
k <- rexp(N,0.5)
beta <- rexp(N,1)
```
```{r}
length(k)
```
not sure why the below not working
```{r}
plot( NULL, xlim=range(d$height),ylim=c(0,500),xlab="height" , ylab="weight" )
#abline( h=0, lty=2 )
abline( h=300 , lty=1 , lwd=0.5 )
mtext( "p ~ rbeta(2,18) & k ~ rexp(0.5) & beta ~ rexp(1)" )
for ( i in 1:N ) curve( 3.141593*k*p^2*x^beta,from=min(d$height) , to=max(d$height) , add=TRUE ,
col=col.alpha("black",0.2) )
```

Looking at other prior info
```{r}
curve( dbeta(x,2,18) )
```
```{r}
curve( dexp(x,0.5) )
```
```{r}
curve( dexp(x,1))
```

```{r}
range(d$h)
```
```{r}
dens(d$h)
```

```{r}
range(d$height)
```