---
title: "Germination_0830"
author: "Lin_Zhang"
date: "8/30/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(dplyr)
library(reshape2)
library(ggplot2)
library(purrr)
```
```{r}
d <- read.csv(file = "light_round1_tall.csv",header = T, sep = ",", stringsAsFactors = F)
```
1.  Subset the data to only include wps==0.  We may look at water potential in a future week.

2. Let's start with a single pop "STDI".  
```{r}
d_wps0 <- filter(d,wps==0)
```
```{r}
d_wps0_STDI <- filter(d_wps0,pops=="STDI")
```
```{r}
ggplot(data = d_wps0_STDI,aes(x = day, y = cumulative_prop_germ))+
  geom_point()+
  geom_line()+
  facet_wrap(~temps)
```
Evaluate models relate germination rate to temperature.  Explain your model choice and prior choice.  Try to do this both with ulam and brm.

```{r}
dat <- list(G = d_wps0_STDI$cumulative_prop_germ,day=d_wps0_STDI$day,germ=d_wps0_STDI$germ, temps=d_wps0_STDI$temps)
```

```{r}
# zero-inflated model as seed under dormancy won't germinate
# censored model as observation window might be shorter than time-range required by seed germination
m1 <- ulam(
alist(
G ~ dzipois( p , lambda ),  # should the dataframe be re-organized thus the maximum germination prop ~ temp used?
logit(p) <- ap + bd*temps,

day|germ > 0 ~ exponential(lambda),
day|germ == 0 ~ custom(exponential_lccdf(!Y| lambda)),
log(lambda) <- ag+bg*temps,

ap ~ dnorm( 0 , 1 ),
ag ~ dnorm( 1 , 0.5 ),
bd ~ dnorm(0,.5),
bg ~ dnorm(0,.5)
) , data=dat , chains=4 )
```
```{r}
plot(precis(m1))
```
```{r}
precis(m1)
```
```{r}
inv_logit(-1.25) # temperature effect on dormancy intercept
inv_logit(0.48) 
exp(-2.74) # temperature effect on germination intercept
exp(-0.06)
```
```{r}
postcheck(m1)
```

3. now expand the data set to all species/populations.  Include species/pop (use "pops" as the predictor, don't worry about having a separate predictor for pops nested within species).  is partial pooling appropriate here?
```{r}
d_wps0$pop_ID <- as.integer(as.factor(d_wps0$pops))
dat <- list(G = d_wps0$cumulative_prop_germ,day=d_wps0$day,germ=d_wps0$germ, temps=d_wps0$temps,Pop=d_wps0$pop_ID)
```

```{r}
# zero-inflated model as seed under dormancy won't germinate
# censored model as observation window might be shorter than time-range required by seed germination
m2 <- ulam(
alist(
G ~ dzipois( p , lambda ),  #dgampois()? how to integrate with dzipois, or partial pooling
logit(p) <- ap[Pop] + bd[Pop]*temps,

day|germ > 0 ~ exponential(lambda),
day|germ == 0 ~ custom(exponential_lccdf(!Y| lambda)),
log(lambda) <- ag[Pop]+bg[Pop]*temps,

ap[Pop] ~ dnorm( -1.5 , 1 ),
ag[Pop] ~ dnorm( 1 , 0.5 ),
bd[Pop] ~ dnorm(0,.5),
bg[Pop] ~ dnorm(0,.5)
) , data=dat , chains=4,log_lik = T)
```
```{r}
precis(m2,depth=2)
```

