---
title: "07_06_20_Phylogenetic_Regression"
author: "Lin_Zhang"
date: "7/6/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(rethinking)
library(ape)
library(ade4)
```

```{r}
data("carni70")
```

# plot the tree
```{r,fig.height=15,fig.width=15}
  carni70.phy <- newick2phylog(carni70$tre)
  plot(carni70.phy, font=1, no.margin=TRUE,label.offset=1 , cex=0.5 )
```
# Model without Phylogenetic Regression
You want to test the hypothesis that the range size of carnivores is influenced by their body size (larger carnivores roam further).  You want to be sure that if you do see a relationship that is is not just due to shared evolutionary history.  That is you want to do a phylogenetic regression.

```{r}
library(data.table)
d <- carni70$tab
setDT(d,keep.rownames = TRUE)
dstan <- d[complete.cases(d$rn,d$size,d$range),]
spp_obs <- dstan$rn
```

3) Should either range or size be transformed?
I guess so 
```{r}
hist(dstan$range)
hist(log(dstan$range))
hist(standardize(log(dstan$range))) #?

hist(dstan$size)
hist(log(dstan$size))
hist(standardize(log(dstan$size)))
```

```{r}
dat_list <- list(
  N_spp = nrow(dstan),
  R = standardize(log(dstan$range)),
  S = standardize(log(dstan$size)),
  Imat = diag(nrow(dstan))
)

# without phylogenetic regression
m1 <- ulam(
  alist(
    R ~ multi_normal(mu, SIGMA),
    mu <- a + bS*S,
    matrix[N_spp,N_spp]:SIGMA <- Imat * sigma_sq,
    a ~ normal(0,1),
    bS ~ normal(0,0.5),
    sigma_sq ~ exponential(1)
      ),data = dat_list, chains = 4, cores = 4)
```
```{r}
precis(m1)
```
looks like a positive association between size and range. But these data are swirling with confounds.

#Phylogenetic Regression
4) Does size predict range?  Evaluate this question without and with accounting for phylogenetic relationships.
2) Note that the species name given in "rownames(d)" and in "tre$tip.label" do not match.  Fix that (e.g. using str_replace() )
handle it when doing Phylogenetic Regression
?
```{r}
#library("stringr") 
#d$rn tree$tip.label look identical to me
#str_replace(d$rn,"[aeiou]",replacement = tree$tip.label)
```
## Brownian motion
```{r}
library(ape)
tree <- as.phylo(carni70.phy)
Rbm <- corBrownian(phy = tree)
V <- vcv(Rbm)
Dmat <- cophenetic(tree)
plot(Dmat, V, xlab="phylogenetic distance", ylab="covariance")
```
```{r}
# put species in right order
dat_list$V <- V[spp_obs,spp_obs]
# convert to correlation matrix 
dat_list$R <- dat_list$V/max(V)

# Brownian motion model
m2 <- ulam(
  alist(
    R ~ multi_normal(mu, SIGMA),
    mu <- a + bS*S,
    matrix[N_spp,N_spp]:SIGMA <- R * sigma_sq,
    a ~ normal(0,1),
    bS ~ normal(0,0.5),
    sigma_sq ~ exponential(1)
  ), data = dat_list, chains = 4, cores = 4
)
precis(m2)
```
the correlation disappeared, the model is suggesting the previous correlation observed is due to phylogenetic relatedness


## OU process
```{r}
# add scaled and reordered distance matrix
dat_list$Dmat <- Dmat[spp_obs,spp_obs]/max(Dmat)

m3 <- ulam(
  alist(
    R ~ multi_normal(mu,SIGMA),
    mu <- a + bS*S,
    matrix[N_spp,N_spp]:SIGMA <- cov_GPL1(Dmat,etasq,rhosq,0.01),
    a ~ normal(0,1),
    bS ~ normal(0,0.5),
    etasq ~ half_normal(1,0.25),
    rhosq ~ half_normal(3,0.25)
  ), data = dat_list, chains = 4, cores = 4
)

precis(m3)
```
There is no correlation between body size and range using this phylogenetic model as well


